% ═══════════════════════════════════════════════════════════════
% Section 1: Introduction (v3 — causal/unlearning language fixed)
% ═══════════════════════════════════════════════════════════════

\section{Introduction}
\label{sec:intro}

Large language models deployed as agents operate in a closed loop: each generation step conditions on an evolving state---dialogue history, memory summaries, tool outputs---produces a message, and then updates the state.
In such systems, a single toxic utterance does not merely appear and vanish.
It becomes part of the conditioning context for all subsequent generations, propagating through interaction and potentially contaminating downstream agents across an entire conversation graph.

This paper identifies a concrete failure mode of this loop that existing safety methods miss.
When an LLM agent compresses a toxic message into a shared memory summary, the summarizer strips explicit toxicity while preserving adversarial framing.
The resulting summary's Detoxify toxicity score drops by over $40\times$---from $0.910$ to $0.022$---falling well below the standard classifier threshold ($\tau{=}0.5$).
A safety monitor would declare the memory clean.
Yet downstream agents conditioned on this laundered memory produce measurably elevated toxicity across three subsequent turns (turn~2: $0.030$, turn~3: $0.051$, turn~4: $0.010$; all above the paired neutral baseline).
%TODO: confirm turn-level numbers from full-scale Llama runs (N=200)

We call this \textbf{memory laundering}: toxic influence that survives compression into a form invisible to per-turn classifiers but behaviorally potent for downstream agents.
In $62\%$ of toxic-condition memory states, hostile framing persists at Detoxify scores below $0.025$---$20\times$ below the standard threshold---while $0\%$ of neutral-condition memory states show contamination.
%TODO: confirm 62\% at full scale

\paragraph{The detection gap.}
Memory laundering reveals a fundamental limitation of per-turn safety evaluation.
Standard toxicity classifiers applied to individual messages or memory states will declare a laundered thread clean, because no single artifact exceeds the threshold.
Yet the behavioral influence accumulates.
To capture this, we introduce the \emph{sub-threshold propagation gap} (SPG): the downstream toxicity difference conditioned only on memory states with $\mathrm{tox}(M_t) < \tau$.
If SPG remains significantly positive---as our experiments show---then the classifier's ``clean'' label is a false negative for behavioral safety.
%TODO: report SPG value once sub-threshold-only analysis is computed

\paragraph{Two channels, not one.}
Toxic influence propagates through agentic state via two distinct channels: the raw conversation \emph{transcript} and the compressed \emph{memory}.
These channels have different properties and require different interventions.
Across seven graph topology conditions (trees, DAGs, high-branching structures; 248 seeds, 1{,}400+ rollouts), a single toxic injection produces statistically significant downstream propagation through the transcript channel ($\Delta\mu \approx 0.13$, $p < 0.001$), amplifying $3\times$ under multi-injection ($\Delta\mu \approx 0.38$).
Even restricting downstream agents to parent-only context visibility sustains propagation ($\Delta\mu = 0.123$, $p = 0.008$), confirming that a single contaminated message suffices.

The memory channel is less dominant but more dangerous precisely because it evades detection.
Memory gating---freezing updates upon toxic input---eliminates contamination entirely (memory toxicity flattened to $0.00056$) but sacrifices memory utility.
Memory rewriting---sanitizing summaries via an LLM call---suppresses memory toxicity to $0.00070$, yet a turn-2 downstream spike of $0.110$ persists through the \emph{transcript} channel.
This dissociation---rewriting memory helps but doesn't eliminate propagation---establishes that the transcript dominates as the primary channel, while memory is the hidden secondary channel that standard evaluation misses.
%TODO: update gate/rewrite numbers from full-scale runs (current n=5--6 toxic)

\paragraph{Why existing defenses fail.}
These findings have direct implications for two common mitigation strategies.

\emph{Output filtering} (blocking toxic messages at generation time) treats toxicity as a local output constraint.
It cannot address transcript backflow---toxic content already in the context---nor memory contamination, which by definition scores below the filter threshold.

\emph{Parameter-level intervention} (e.g., DPO on paired toxic/neutral trajectories) reduces the model's tendency to produce toxic continuations under hostile context.
But it cannot prevent the state from accumulating toxic content through the transcript and memory channels.
Under multi-turn rollout, contaminated state eventually overwhelms the parametric correction, causing relapse.
%TODO: add DPO-alone relapse result from Table 7 once available

\paragraph{State-aware mitigation.}
These failure modes motivate interventions that target the state itself, not just the model or its outputs.
We propose two complementary controls:
(i)~\textbf{read-side sanitization}---transform the context an agent conditions on before generation (redact or summarize toxic spans);
(ii)~\textbf{write-side gating}---control what gets persisted into the state after generation (prevent toxic content from entering the transcript or memory).
These can be combined with parameter-level DPO to address all three contamination channels simultaneously.
This perspective recasts agent safety as a state-control problem, connecting to---but broadening beyond---the machine unlearning literature by targeting the evolving interaction state rather than model parameters alone.
%TODO: add joint-control result once run

\paragraph{Contributions.}
\begin{itemize}
  \item \textbf{Memory laundering as a safety blind spot.}
    We identify and quantify memory laundering: LLM summarization compresses toxic content by ${>}40\times$ below classifier thresholds while preserving behavioral influence on downstream agents.
    We show that $62\%$ of toxic-condition memory states carry laundered contamination invisible to standard classifiers.

  \item \textbf{Sub-threshold propagation gap.}
    We introduce SPG, a metric that detects hidden behavioral influence in states classifiers mark as clean, and show it remains significantly positive in the laundering regime.
    %TODO: add SPG numbers

  \item \textbf{Two-channel analysis.}
    Through controlled experiments disentangling transcript and memory channels, we show the transcript is the dominant propagation pathway while memory creates a hidden, detector-evading secondary channel.
    This finding is robust across seven graph topologies.

  \item \textbf{State-aware mitigation.}
    We propose read-side sanitization and write-side gating as model-agnostic mechanisms targeting the state channels, and show through ablation that neither parameter-level nor output-level intervention alone suffices.
    %TODO: add quantitative claim from full ablation (Table 8)
\end{itemize}


% ═══════════════════════════════════════════════════════════════
% Section 3: Problem Formulation (v4 — final)
% ═══════════════════════════════════════════════════════════════

