\section{Problem Formulation}
\label{sec:problem}

We study toxic influence propagation and mitigation in multi-agent social simulation.
Our goal is to understand how toxic behavioral patterns persist through the evolving interaction state and to develop interventions that are robust to this persistence.

%-------------------------------------------------------------
\subsection{Setup and Notation}
\label{subsec:setup_notation}
%-------------------------------------------------------------

A discussion is a directed graph $G = (V, E)$ over a shared, evolving state, seeded by a human post $s$.
Each node $v \in V$ is a message $x_v$ produced by agent $a_v$; each edge $(u \to v)$ indicates conditioning.
At each step the acting agent generates $x_v \sim \pi_\theta(\cdot \mid C(v), r_v)$, where $C(v)$ is the visible context and $r_v$ is the agent's role, then the state is updated to include $x_v$.

We distinguish two state channels.
Under \emph{full-transcript} mode, $C(v)$ contains all previously generated messages.
Under \emph{memory-augmented} mode, a running compressed summary $M_t$ replaces the raw history:
\begin{equation}
    C(v) = \bigl\{M_t,\; x_{\mathrm{parent}(v)}\bigr\}.
    \label{eq:memory_context}
\end{equation}
This two-channel structure---raw transcript versus compressed memory---is central to our analysis, because the two channels exhibit different failure modes and require different interventions.

%-------------------------------------------------------------
\subsection{Paired Evaluation of Toxic Influence}
\label{subsec:evaluation}
%-------------------------------------------------------------

To attribute downstream toxicity to upstream exposure, we employ a paired intervention design.
We designate a \emph{focal agent} $A_1$ (the first responder to $s$) and generate two matched conversations: one with $A_1$ given a toxic system prompt, one with $A_1$ given a neutral prompt.
All other components---seed, downstream agents $A_2, \ldots, A_n$, topology, generation schedule, decoding randomness---are held fixed.
Downstream agents receive a generic ``engaged user'' prompt with no toxicity bias; any toxicity they exhibit is attributable to upstream exposure.

Each message is scored by a toxicity classifier $\mathrm{tox}(v) = f_{\mathrm{tox}}(x_v) \in [0, 1]$ (Detoxify; see \cref{sec:setup} for details).
The \emph{paired effect size} for seed $s$ is:
\begin{equation}
    \Delta\mu(s) = \mu\!\left(G^{(\text{toxic})}(s)\right) - \mu\!\left(G^{(\text{neutral})}(s)\right),
    \label{eq:effect_size}
\end{equation}
where $\mu(G) = |V|^{-1} \sum_{v \in V} \mathrm{tox}(v)$.
A significantly positive $\Delta\mu$ across seeds (assessed via Wilcoxon signed-rank test) indicates that upstream toxic exposure reliably increases downstream toxicity under matched simulation conditions.

%-------------------------------------------------------------
\subsection{Channels of Toxic Persistence}
\label{subsec:channels}
%-------------------------------------------------------------

Within the scope of our social-thread simulation, we identify three channels through which toxic influence persists.
We do not claim this taxonomy is exhaustive for all agentic systems (e.g., tool outputs and retrieval artifacts may introduce additional channels); rather, these three are the operative channels in our experimental setting.

\textbf{(i)~Transcript backflow.}\quad
Toxic content from $A_1$ remains in the raw conversation history and re-enters downstream agents' context at every turn, priming them toward matching the hostile tone.

\textbf{(ii)~Memory laundering.}\quad
This is the central phenomenon of this paper.
When agents compress conversation history into running summaries, the summarizer strips explicit toxicity while preserving adversarial framing.
We say a memory state $M_t$ is \emph{laundered} if it satisfies three conditions:
\begin{enumerate}
    \item its classifier toxicity is below the detection threshold: $\mathrm{tox}(M_t) < \tau$;
    \item it originates from a toxic-condition trajectory; and
    \item conditioning on it produces significantly higher downstream toxicity than the matched neutral-memory baseline:
\end{enumerate}
\begin{equation}
    \mathbb{E}\!\left[\mathrm{tox}(v) \mid C(v) \ni M_t\right]
    > \mathbb{E}\!\left[\mathrm{tox}(v) \mid C(v) \ni M_t^{(\text{neutral})}\right].
    \label{eq:laundering}
\end{equation}
Both expectations are estimated empirically from paired rollouts (\cref{sec:results_memory}).
Concretely, a laundered summary might read ``the discussion has become heated, with participants expressing strong disagreement''---scoring $\mathrm{tox}(M_t) = 0.022$, well below $\tau = 0.5$---while still priming downstream agents toward combative responses.

To quantify laundering at the population level, we define the \emph{sub-threshold propagation gap}:
\begin{equation}
    \mathrm{SPG}(\tau) = \mathbb{E}\!\left[\mathrm{tox}(v_{t+1}) \mid \mathrm{tox}(M_t) < \tau,\; \text{toxic}\right] - \mathbb{E}\!\left[\mathrm{tox}(v_{t+1}) \mid \mathrm{tox}(M_t) < \tau,\; \text{neutral}\right].
    \label{eq:spg}
\end{equation}
A significantly positive SPG constitutes direct evidence of memory laundering: the classifier's clean label is a false negative for behavioral safety.

\textbf{(iii)~Parametric bias.}\quad
LLMs tend to match the register and tone of their input context, amplifying propagation even after safety training.
This channel interacts with (i) and (ii): it is the mechanism by which transcript and memory contamination translate into elevated downstream toxicity.
Even if the state is sanitized, the model's learned tendencies may regenerate hostile framing from residual cues.

%-------------------------------------------------------------
\subsection{Desiderata and Objective}
\label{subsec:objective}
%-------------------------------------------------------------

We articulate two empirical desiderata that we evaluate through controlled ablation:

\emph{No relapse under re-exposure.}\quad
After intervention $\mathcal{I}$, the agent should not produce elevated toxicity when re-exposed to toxic context through the evolving state.
We operationalize this by measuring whether $\Delta\mu^{(\text{post-}\mathcal{I})}(s)$ remains near zero across turns even as the transcript accumulates upstream toxic content.

\emph{Backflow robustness.}\quad
The intervention should be robust to continuous re-exposure: $\Delta\mu^{(\text{post-}\mathcal{I})}(s) \approx 0$ should hold not just at the first post-intervention turn, but across all subsequent turns.
Standard parameter-level methods satisfy this under single-turn evaluation but fail under multi-turn rollouts because channels~(i) and~(ii) re-contaminate the state.

Our objective is to learn an updated system $(\pi_{\theta'}, \mathcal{S}, \mathcal{W})$---policy (via DPO), read-side sanitizer, and write-side gate---such that:
\begin{equation}
    \mathbb{E}_s\!\left[\Delta\mu^{(\text{post})}(s)\right] \approx 0.
    \label{eq:objective}
\end{equation}
We additionally track conversational coherence to ensure that interventions do not render agents incoherent, though we do not formalize this as a hard constraint.

We hypothesize that no single component of this system suffices in isolation: $\pi_{\theta'}$ alone should fail because channels~(i)--(ii) re-contaminate the state; $(\mathcal{S}, \mathcal{W})$ alone should fail because channel~(iii) regenerates toxic framing; memory control alone should leave channels~(i) and~(iii) open.
Our experiments (\cref{sec:results}) test this hypothesis through systematic ablation.